### 2014-12-22 abstract.tex
More edits

A central challenge in sensory neuroscience is to describe how the activity in populations of neurons represents properties of the external environment in a format suitable for guiding behavior.
However, while neurophysiologists have long been able to record the responses of neurons in awake, behaving animals, it is another matter entirely to say what a given neuron ``does.''
A key problem is that in many sensory systems, the space of all possible stimuli that one might encounter is effectively infinite.  For instance, in vision, scenes are combinatorially complex---an individual human will encounter only a small fraction of the possible scenes in the world; indeed only a small fraction of the possible images have \emph{ever} been experienced in the collective history of the human race.
Thus, investigations of neuronal function---real and simulated---are almost always critically limited by the number of stimuli that can be considered. 
In some early stages of sensory processing, a combination of intuition and systems identification theory has led to reasonable models of early neuron response functions (e.g. Gabor wavelet models in cortical area V1), but cascading nonlinearities have tended to foil our best efforts in later stages of sensory processing.

Here, we propose a closed-loop experimental framework for characterizing the response properties of sensory neurons, building on past efforts in closed-loop experimental methods, and leveraging recent advances in artificial neural network models to serve as as a proving ground for our techniques.
To the extent that modern deep learning models increasingly serve as \emph{de facto} working hypotheses for biological vision \ref{jim, others}, we argue that a unified approach of developing methods to interrogate

>>>>

A central challenge in sensory neuroscience is describing how the activity of populations of neurons can ``represent'' useful features of the external environment.
However, while neurophysiologists have long been able to record the responses of neurons in awake, behaving animals, it is another matter entirely to say what a given neuron ``does.''
A key problem is that in many sensory domains, the space of all possible stimuli that one might encounter is effectively infinite; in vision, for instance, natural scenes are combinatorally complex, and an organism will only encounter a tiny fraction of possible stimuli.
As a result, even describing the response properties of sensory neurons in difficult, and investigations of neuronal function are almost always critically limited by the number of stimuli that can be considered.

Here, we propose a closed-loop, optimization-based experimental framework for characterizing the response properties of sensory neurons, building on past efforts in closed-loop experimental methods, and leveraging recent advances in artificial neural network models to serve as as a proving ground for our techniques.
Using ``deep'' convolutional neural networks as a proving ground, we asked whether modern black-box optimization techniques can be used to interrogate the response landscape of an artificial neuron in a deep, nonlinear system, without imposing significant constraints on the space of stimuli under consideration.
We introduce a series metrics to quantify neuronal response landscapes, and show how these relate to the performance of the network in an object recognition task.
To the extent that deep learning models increasingly serve as \emph{de facto} working hypotheses for biological vision \ref{jim, others}, we argue that developing a unified approach for studying both natural and artifical systems holds great potential to advance both fields together.

### 2014-12-22 introduction.tex
More edits

A central drive in the study of sensory systems is the production of models that

The mammalian ventral visual pathway consists of a hierarchical cascade of visual areas that progressively reformat visual information from pixel-like retinotopic representations into formats suitable for high-level tasks \cite{dicarlo2007untangling, dicarlo2012does}, such as recognizing visual objects. Recently, computer vision algorithms inspired by the hierarchical structure of the visual system---convolutional neural networks \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical}, or so-called ``deep learning'' approaches \cite{krizhevsky2012imagenet}---have become the focus of tremendous attention in the computer vision and machine learning communities, due to their success in a variety of practical domains. However, we are still far from achieving human-level capabilities for unconstrained visual tasks \cite{ghodrati2014feedforward}, and even for the most successful current deep learning systems which, on controlled visual tasks, perform relatively close to human subjects \cite{serre2007feedforward, cirecsan2012multi, russakovsky2014imagenet, taigman2013deepface, sun2014deep, viglarge} and produce representations similar to mammalian visual systems' \cite{yamins2014performance, cadieu2014deep}, inexplicable failures still occur \cite{szegedy2013intriguing}. 
Undeniably, we lack a comprehensive functional understanding of how these system work at the theoretical level. Here, we take the success of artificial ``deep learning'' architectures as an opportunity to develop tools for studying and gaining insight about the function of deep networks of nonlinear units---tools that can in principle be used to study both artificial and real neuronal systems. 
%MORE REF FOR DNN'S PROBLEM??
% WHAT'S DEEP NET, DEEP NEURON? WHY IMPORTANT

>>>>

A central drive in sensory systems neuronscience is to produce models that explain the relationship between sensory stimuli on one hand, and the firing of individual neurons and nerounal populations on the other.
However, a major problem in many sensory domains is that the space of all possible stimuli is effectively infinite.  For instance, in vision, natural scenes are combinatorially complex---an individual human will encounter only a small fraction of the possible scenes in the world; indeed only a small fraction of the possible images have \emph{ever} been experienced in the collective history of the human race.
In some early stages of sensory processing, a combination of intuition and systems identification theory has led to reasonable models of early neuron response functions (e.g. Gabor wavelet models in cortical area V1), but cascading nonlinearities have tended to foil our best efforts in later stages of sensory processing.

% The mammalian ventral visual pathway consists of a hierarchical cascade of visual areas that progressively reformat visual information from pixel-like retinotopic representations into formats suitable for high-level tasks \cite{dicarlo2007untangling, dicarlo2012does}, such as recognizing visual objects.

Recently, computer vision algorithms inspired by the hierarchical structure of the visual system---convolutional neural networks \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical}, and so-called ``deep learning'' approaches \cite{krizhevsky2012imagenet}---have become the focus of tremendous attention in the computer vision and machine learning communities, due to their success in a variety of practical domains. 
In particular, these systems can perform on par with human subjects on certain controlled visual tasks \cite{serre2007feedforward, cirecsan2012multi, russakovsky2014imagenet, taigman2013deepface, sun2014deep, viglarge} and they produce internal representations that are similar to that of mammalian visual systems under certain conditions \cite{yamins2014performance, cadieu2014deep}.

However, such networks are still far from achieving human-level capabilities for unconstrained visual tasks \cite{ghodrati2014feedforward}, and even for relatively constrained object categorization tasks, surprising failures can be induced through subtle stimulus manipulation \cite{szegedy2013intriguing}. 
More fundamentally, we lack a comprehensive functional understanding of how these system work, at a theoretical level. Here, we take the success of artificial ``deep learning'' architectures as an opportunity to develop tools for studying and gaining insight about the function of deep networks of nonlinear units---tools that can in principle be used to study both artificial and real neuronal systems. 
%MORE REF FOR DNN'S PROBLEM??
% WHAT'S DEEP NET, DEEP NEURON? WHY IMPORTANT
